Symmetric Peer-to-Peer Applications

### 2022-08-28 tpd-22.tex
...

% docs.google.com/spreadsheets/d/1CpMmEgabJq2XQeTDW_BNCV4JWrcMlELuaEjDtw7kwg0/
\begin{figure}[t]
  \centering
  \includegraphics[width=\linewidth]{table}
  \caption{
    \label{fig.table}
  }
\end{figure}

@@

The middleware assumes that each peer is assigned a unique identifier in a
contiguous range.
The FPS must be the same in all peers to ensure that the simulation executes
exactly the same.

@@

\subsection{TODO}

TODO: simpler than vector clocks, absolute times are correct
TODO: piggyback SYNC in normal inputs

@@

We now describe in detail how the middleware orchestrates an application and
ensures that all peers (a) meet event deadlines, (b) advance in sync in real
time, and (c) can leave and join the network and remain symmetric.

@@

%\item Peers have contiguous numeric unique identifiers.

@@

We make some assumptions about the network, which are assigned externally to
the middleware:
%
\begin{itemize}
\item Peers have contiguous numeric unique identifiers.
\item Peers identifiers are assigned externally to the middleware.
\item Topology is dynamic, but also assigned externally

@@

, which we claim that could be
solved further:
- peers have unique identifiers
- ids are outside
- topology, dynamic but outside
- fixed maximum nodes
- fixed latency
We claim that Use external protocols
For instance, latency is only a constant, no problem being a variable

>>>>

% docs.google.com/spreadsheets/d/1CpMmEgabJq2XQeTDW_BNCV4JWrcMlELuaEjDtw7kwg0/
\begin{figure}[t]
  \centering
  \includegraphics[width=\linewidth]{table}
  \caption{
    Related work regarding
        (a) how the network is organized,
        (b) how global time is determined, and
        (c) how events are propagated and applied.
    \label{fig.table}
  }
\end{figure}

@@

The middleware assumes that each peer assigns to itself a unique identifier in
a contiguous range.
The FPS must be the same in all peers to ensure that the simulation executes
exactly the same.

@@

@@

We now detail how the middleware orchestrates an application and ensures that
all peers
    (a) meet event deadlines,
    (b) advance in sync in real time, and
    (c) can leave and join the network and remain symmetric.

\subsubsection{Events Broadcasting}
\label{sec.tml.middleware.events}

As illustrated in Figure~\ref{fig.middleware}, events are broadcast between
peers with a timestamp in the future such that all peers are able to apply them
in sync.
%
To prevent broadcast cycles, event are tagged with a source peer and sequence
number.
Each peer keeps a vector with the highest sequence number it received from each
other peer, such that lower numbers can be ignored when received.
Higher numbers update the vector and trigger a broadcast to neighbours.
Since we rely on TCP connections, it is not possible to receive out-of-order
sequence numbers from the same source peer.
%
As mentioned in Section~\ref{sec.tml.api}, we assume that peers assign
themselves unique identifiers in a contiguous range.
The reasons are
    (a) to fit a \code{uint8\_t} size, and
    (b) to use a contiguous vector of sequence numbers.
Therefore, this is not a serious limitation, and a separate mechanism could be
used for assignment, with few modifications to core of the middleware.

When receiving an unseen event, the middleware also enqueues it in timestamp
order and calls the application callback \code{cb\_sim} at the appropriate
time.
The peers must keep the full queue in memory for two reasons:
    (a) receiving a late event may reorder the queue and require a time
        travel, and
    (b) peers that join the network need to receive all events.
%
Therefore, the middleware keeps a queue of ``packets'', which includes the
event and other metadata:

{\footnotesize
\begin{verbatim}
typedef struct {
    uint8_t  src;       // source peer
    uint32_t seq;       // sequence number
    uint32_t tick;      // tick to apply
    p2p_evt  evt;       // event id + payload
} p2p_pak;

struct {
    int i;              // next event to apply
    int n;              // number of events
    p2p_pak buf[MAX];   // queue of packets
} PAKS;
\end{verbatim}
}

\subsubsection{The Main Loop}
\label{sec.tml.middleware.loop}

The middleware main loop \code{p2p\_loop} is responsible for calling the
application callbacks, and also controlling its timeline.
To achieve this, the loop continuously checks for network packets, local
inputs, and time ticks, as follows:

{\footnotesize
\begin{verbatim}
01  void p2p_loop (...,cb_ini,cb_sim,cb_ren,cb_evt) {
01    cb_ini(1);
01    cb_sim(<P2P_EVT_INIT>);
01
01    while (<app-running>) {
01      if (<next-packet>) {
01        // fast forward if too late
01        if (<packet-sync-time-future>) {
01          for (<step-by-step-sleep>) {
01            p2p_travel(<packet-sync-time>);
01            cb_ren();
01          }
01        }
01
01        if (<packet-evt>) {
01          if (<packet-evt-time-past>) {
01            for (<step-by-step-sleep>) {
01              p2p_travel(<packet-evt-time>);
01              cb_ren();
01            }
01            for (<step-by-step-sleep>) {
01              p2p_travel(<now>);
01              cb_ren();
01            }
01          } else {
01            cb_sim(<packet-evt>);
01            cb_eff();
01          }
01        }
01      }
01
01      if (<next-input>) {
01        p2p_evt evt;
01        if (cb_rec(&evt)) {
01          p2p_bcast(<future>, &evt);
01        }
01      }
01
01      if (<next-tick>) {
01        cb_sim(<P2P_EVT_TICK>);
01        <take-snapshot-every-XXXms>
01        cb_ren();
01      }
01    }
01
01    cb_ini(0);
01  }
\end{verbatim}
}

- async events
- sync events
- real-time sim
- sync time

advance time locally

TODO: no final falar de a,b,c
TODO: simpler than vector clocks, absolute times are correct
TODO: piggyback SYNC in normal inputs

@@

@@

We make some assumptions about the network, which are assigned externally to
the middleware:
%
\begin{itemize}
\item Peers identifiers are assigned externally to the middleware.
\item Topology is dynamic, but also assigned externally

@@

, which we claim that could be
solved further:
- ids are outside
- topology, dynamic but outside
- fixed maximum nodes
- fixed latency
We claim that Use external protocols
For instance, latency is only a constant, no problem being a variable
